Box

### 2021-07-17 MS/box.Rmd
boxed mended with a few new refs

Here we show how to obtain sampling variance for a slope (or a 'rate of change') when you have measurements of a physiological trait at two points along an environmental gradient (e.g. temperature, salinity, acidity). We also go on to derive sampling variance for the difference between two slopes, and finally end on showing how SMD can be corrected to account for both unitary and temperature differences across effect sizes. The sampling variances associated with these 'new' effect sizes will allow one to take advantage of powerful meta-analytic models.

@@

To obtain the sampling variance for this equation (slope), we need to know several basic properties of variance. Let's assume $M_{1}$ is a random variable, which means that this variable has a mean and standard deviation (variance). Multiplying a constant ($a$) will change the variance by the square of that constant ($a^2$) while adding or subtracting the constant ($b$) does not change the variance of $M_{1}$. This can be summarized as:

@@

$$
\sigma^2_{M_1 \pm M_2} = \sigma^2_{M_1} + \sigma^2_{M_2} \pm 2\text{Cov}(M_1,M_2) 
= \sigma^2_{M_1} + \sigma^2_{M_2} \pm 2\text{Cor}(M_1,M_2)\sqrt{ \sigma^2_{M_1} \sigma^2_{M_2} }
$$ 
where the co-variance $\text{Cov}(M_1,M_2)$ equals to the correlation multiplied by the square-root of the two variances $2\text{Cor}(M_1,M_2)\sqrt{ \sigma^2_{M_1} \sigma^2_{M_2} }$.

@@

where $SD_{1}$ and $SD_{2}$ and $N_{1}$ and $N_{2}$ are standard deviations and sampling sizes at temperature $T_{1}$ and $T_{2}$. Readers may still find it difficult to see how we got this equation. Let us explain further. The (sampling) standard error for $X_{1}$ is $SE_{1} = SD_{1} / \sqrt{N_{1}}$ -- many readers will probably know this as the formula for obtaining *SE* from *SD*. Given this, the sampling variance for $X_{1}$ is $SD^2_{1} / N_{1}$, which you see as an element of the sampling variance of ARR along with the sampling variance for $X_{2}$, which is $X_{2}$ is $SD^2_{2} / N_{2}$; the term $(1/(T_{2} - T_{1}))^2$ comes from recognizing $1/(T_{2} - T_{1})$ as a constant.

@@

Here, we can see that the difference between $M_1$ and $M_2$ is standardised by the pooled SD, essentially correcting for unitary differences. Additionally, dividing SMD by the temperature difference results in further correcting the effect size by temperature. To derive the sampling variance for $SMD_T$ we can apply the same principles as we did above to derive an approximate sampling variance as:

@@

It is often limited when we can use the basic properties of variance to derive sampling variance. Therefore, we introduce a practical and widely applicable method to obtain approximate variance when the basic properties of variance cannot be applied. The method is widely known as the delta method, which can be written as:

Here, we use the delta method to derive the sampling variance for the log repose ratio, which is:

@@

Finally, we note that by using the basic properties of variance and the delta method, one can derive sampling variance for most of cases. Indeed, these were enough for us to derive sampling variance for 'new' effect sizes (e.g. $lnQ_{10}$; see the main text) and as in @Nakagawa2015 and @Senior2020.

>>>>

Here we show how to obtain sampling variance for a slope (or a 'rate of change') when you have measurements of a physiological trait at two points along an environmental gradient (e.g. temperature, salinity, acidity). We then go on to derive sampling variance for the difference between two slopes, and also how SMD (standardized mean difference) can be corrected to account for differences in both units (e.g. cm or mg) and points on an environmental gradient (e.g. temperature of 24 and 30 Celsius). Finally, we introduce a useful approximation technique known as 'the delta method'. The sampling variances associated with these 'new' effect sizes below will allow one to take advantage of powerful meta-analytic models.

@@

To obtain the sampling variance for this equation (slope), we need to know several basic properties of variance. Let's assume $M_{1}$ is a random variable, which means that this variable has a mean and standard deviation (variance; this standard decision is not 'sample' but 'sampling' standard deviation, which is often referred to as standard error). Multiplying a constant ($a$) will change the variance by the square of that constant ($a^2$) while adding or subtracting the constant ($b$) does not change the variance of $M_{1}$. This can be summarized as:

@@

$$
\sigma^2(M_1 \pm M_2) = \sigma^2_{M_1} + \sigma^2_{M_2} \pm 2\text{Cov}(M_1,M_2) 
= \sigma^2_{M_1} + \sigma^2_{M_2} \pm 2\text{Cor}(M_1,M_2)\sqrt{ \sigma^2_{M_1} \sigma^2_{M_2} }
$$

where the co-variance $\text{Cov}(M_1,M_2)$ equals to the correlation multiplied by the square-root of the two variances $2\text{Cor}(M_1,M_2)\sqrt{ \sigma^2_{M_1} \sigma^2_{M_2} }$.

@@

where $SD_{1}$ and $SD_{2}$ and $N_{1}$ and $N_{2}$ are standard deviations and sampling sizes at temperature $T_{1}$ and $T_{2}$. Readers may still find it difficult to see how we got this equation. Let us explain further. The (sampling) standard error for $X_{1}$ is $SE_{1} = SD_{1} / \sqrt{N_{1}}$ -- many readers will probably know this as the formula for obtaining *SE* ('sampling' standard deviation) from *SD* ('sample' standard deviation; see Figure 1 in @Nakagawa2021b). Given this, the sampling variance for $X_{1}$ is $SD^2_{1} / N_{1}$, which you see as an element of the sampling variance of ARR along with the sampling variance for $X_{2}$, which is $X_{2}$ is $SD^2_{2} / N_{2}$; the term $(1/(T_{2} - T_{1}))^2$ comes from recognizing $1/(T_{2} - T_{1})$ as a constant.

@@

Here, we can see that the difference between $M_1$ and $M_2$ is standardised by the pooled SD, essentially correcting for unitary differences. Additionally, dividing SMD by the temperature difference results in further correcting the effect size by temperature. To derive the sampling variance for $SMD_T$ we can apply the same principle as we did above to derive an sampling variance as:

@@

It is often limited when we can use the basic properties of variance to derive sampling variance. Therefore, we introduce a practical and widely applicable method to obtain approximate variance when the basic properties of variance cannot be applied. This is where the delta method comes in. The general form of this method can be written as:

@@

Finally, we note that by using the basic properties of variance and the delta method, one can derive sampling variance for most of cases. Indeed, these were enough for us to derive sampling variance for 'new' effect sizes (e.g. $lnQ_{10}$; see the main text) and as in @Nakagawa2015 and @Senior2020. Also, see @Nakagawa2017 for more on the delta method (they explain the 'first and second-order' delta method)
